---
title: "CBE 2"
author: "Amalia, Lauren, and Jenny"
date: last-modified
format:
  pdf:
    number-sections: true
    indent: true
    toc: true
    include-in-header:
      text: |
        \addtokomafont{disposition}{\rmfamily}
        \usepackage{fvextra}
        \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
execute:
  echo: false
---

\newpage

# Introduction

This project explores the differences in writing styles between the 1850s and the 1950s, with a focus on why older texts are often harder to understand. By examining a collection of short stories from both periods, we look at how factors like sentence length and lexical diversity may influence the readability of older texts. This question is interesting because understanding these shifts in writing style can reveal how language and communication have changed over time, and whether these shifts have made texts more accessible or harder to read.

# Data

```{r, include=FALSE}
library(quanteda)
library(quanteda.textstats)
library(tidyverse)
library(sentimentr)
library(knitr)
library(gt)
library(syuzhet)
library(tidytext)
library(dplyr)
source("../R/helper_functions.R")
source("../R/keyness_functions.R")

```

```{r}

txt_1945 <- paste(readLines("Text Files/the_other_side_of_the_hedge_1945.txt"), collapse = " ")

txt_1852 <- paste(readLines("Text Files/the_old_nurses_story_1852.txt"), collapse = " ")

txt_1952 <- paste(readLines("Text Files/the_old_man_at_the_bridge_1952.txt"), collapse = " ")

txt_1950 <- paste(readLines("Text Files/the_mindworm_1950.txt"), collapse = " ")

txt_1842 <- paste(readLines("Text Files/the_masque_of_the_red_death_1842.txt"), collapse = " ")

txt_1940 <- paste(readLines("Text Files/the_man_who_killed_the_world_1940.txt"), collapse = " ")

txt_1859 <- paste(readLines("Text Files/the_lifted_veil_1859.txt"), collapse = " ")

txt_1839 <- paste(readLines("Text Files/the_fall_of_the_house_of_usher_1839.txt"), collapse = " ")

txt_1852 <- paste(readLines("Text Files/the_district_doctor_1852.txt"), collapse = " ")

txt_1954 <- paste(readLines("Text Files/the_destructors_1954.txt"), collapse = " ")

txt_1942 <- paste(readLines("Text Files/the_catbird_seat_1942.txt"), collapse = " ")

txt_1858 <- paste(readLines("Text Files/psyche's_art_1858.txt"), collapse = " ")


txt_1954b <- paste(readLines("Text Files/master_zacharius_1954.txt"), collapse = " ")

txt_1955 <- paste(readLines("Text Files/good_country_people_1955.txt"), collapse = " ")


txt_1853 <- paste(readLines("Text Files/gods_in_exile_1853.txt"), collapse = " ")


txt_1950b <- paste(readLines("Text Files/first_confession_1950.txt"), collapse = " ")

txt_1950c <- paste(readLines("Text Files/coming_attraction_1950.txt"), collapse = " ")

txt_1853b <- paste(readLines("Text Files/bartelby_the_scrivener_1853.txt"), collapse = " ")

txt_1852b <- paste(readLines("Text Files/a_terribly_strange_bed_1852.txt"), collapse = " ")

txt_1952b <- paste(readLines("Text Files/a_sound_of_thunder_1952.txt"), collapse = " ")

txt_1948 <- paste(readLines("Text Files/a_perfect_day_for_bananafish_1948.txt"), collapse = " ")

txt_1951 <- paste(readLines("Text Files/A_&_P_1951.txt"), collapse = " ")

```

```{r,include=FALSE}
years_q <- c("1839", "1842", "1852", "1852b", "1853", "1853b", "1858", "1859", "1940", "1942", "1945", "1948", "1950", "1950b", "1950c", "1951", "1952", "1952b", "1954", "1954b", "1955")

years <- c(txt_1839, txt_1842, txt_1852, txt_1852b, txt_1853, txt_1853b, txt_1858, txt_1859, txt_1940, txt_1942, txt_1945, txt_1948, txt_1950, txt_1950b, txt_1950c, txt_1951, txt_1952, txt_1952b, txt_1954, txt_1954b, txt_1955)

years_18_q <- c("1839", "1842", "1852", "1852b", "1853", "1853b", "1858", "1859")
years_19_q <- c("1940", "1942", "1945", "1948", "1950", "1950b", "1950c", "1951", "1952", "1952b", "1954", "1954b", "1955")

years_18 <- c(txt_1839, txt_1842, txt_1852, txt_1852b, txt_1853, txt_1853b, txt_1858, txt_1859)
years_19 <- c(txt_1940, txt_1942, txt_1945, txt_1948, txt_1950, txt_1950b, txt_1950c, txt_1951, txt_1952, txt_1952b, txt_1954, txt_1954b, txt_1955)


short_stories_corpus <- data.frame(doc_id = years_q, text = years) %>%
   mutate(text = preprocess_text(text)) %>%
   corpus()

short_stories_18_corpus <- data.frame(doc_id = years_18_q, text = years_18) %>%
   mutate(text = preprocess_text(text)) %>%
   corpus()

short_stories_19_corpus <- data.frame(doc_id = years_19_q, text = years_19) %>%
   mutate(text = preprocess_text(text)) %>%
   corpus()

dfm_short_stories <- dfm_remove(dfm(tokens(short_stories_corpus)), stopwords("en"))

dfm_short_stories_18 <- dfm_remove(dfm(tokens(short_stories_18_corpus)), stopwords("en"))

dfm_short_stories_19 <- dfm_remove(dfm(tokens(short_stories_19_corpus)), stopwords("en"))
```

```{r}
tokens_list <- tokens(short_stories_corpus)
tokens_lexical <- tokens_select(tokens_list, pattern = stopwords("en"), selection = "remove")
```

```{r}
calculate_asl <- function(text) {
   unlist(str_split(text, pattern = "\\.\\s*")) %>%
    sapply(function(x) str_count(x, "\\w+")) %>%
    mean()
}


calculate_awl <- function(tokens) {
  tokens %>%
    tokens(what = "word") %>%
    as.list() %>%
    unlist() %>%
    nchar() %>%
    mean()
}

results <- tibble(
  Year = years_q,
  ASL = sapply(years, calculate_asl),
  AWL = sapply(short_stories_corpus, calculate_awl)
)

results %>%
  arrange(Year) %>%
  print()

```

```{r}
all_text <- paste(txt_1839, txt_1842, txt_1852, txt_1852b, txt_1853, txt_1853b, txt_1858, txt_1859)
words <- unlist(strsplit(all_text, "\\s+"))
total_word_count <- length(words)
print(total_word_count)
all_text1950s <- paste(txt_1940, txt_1942, txt_1945, txt_1948, txt_1950, txt_1950b, txt_1950c, txt_1951, txt_1952, txt_1952b, txt_1954, txt_1954b, txt_1955)
words1950s <- unlist(strsplit(all_text1950s, "\\s+"))
total_word_count1950s <- length(words1950s)
```
```{r}
word_count_1839 <- length(unlist(strsplit(txt_1839, "\\s+")))
word_count_1842 <- length(unlist(strsplit(txt_1842, "\\s+")))
word_count_1852 <- length(unlist(strsplit(txt_1852, "\\s+")))
word_count_1852b <- length(unlist(strsplit(txt_1852b, "\\s+")))
word_count_1853 <- length(unlist(strsplit(txt_1853, "\\s+")))
word_count_1853b <- length(unlist(strsplit(txt_1853b, "\\s+")))
word_count_1858 <- length(unlist(strsplit(txt_1858, "\\s+")))
word_count_1859 <- length(unlist(strsplit(txt_1859, "\\s+")))
word_counts1850s <- c(word_count_1839, word_count_1842, word_count_1852, word_count_1852b, word_count_1853, word_count_1853b, word_count_1858, word_count_1859)
average_word_count1850s <- mean(word_counts1850s)

#average word count 1850s: 8718.625
#average word count 1950s: 5379.615
```

```{r}
word_count_1940 <- length(unlist(strsplit(txt_1940, "\\s+")))
word_count_1942 <- length(unlist(strsplit(txt_1942, "\\s+")))
word_count_1945 <- length(unlist(strsplit(txt_1945, "\\s+")))
word_count_1948 <- length(unlist(strsplit(txt_1948, "\\s+")))
word_count_1950 <- length(unlist(strsplit(txt_1950, "\\s+")))
word_count_1950b <- length(unlist(strsplit(txt_1950b, "\\s+")))
word_count_1950c <- length(unlist(strsplit(txt_1950c, "\\s+")))
word_count_1951 <- length(unlist(strsplit(txt_1951, "\\s+")))
word_count_1952 <- length(unlist(strsplit(txt_1952, "\\s+")))
word_count_1952b <- length(unlist(strsplit(txt_1952b, "\\s+")))
word_count_1954 <- length(unlist(strsplit(txt_1954, "\\s+")))
word_count_1954b <- length(unlist(strsplit(txt_1954b, "\\s+")))
word_count_1955 <- length(unlist(strsplit(txt_1955, "\\s+")))

word_counts1950s <- c(word_count_1940, word_count_1942, word_count_1945, word_count_1948, word_count_1950, word_count_1950b, word_count_1950c, 
                      word_count_1951, word_count_1952, word_count_1952b, 
                      word_count_1954, word_count_1954b, word_count_1955)
average_word_count1850s <- mean(word_counts1950s)
```

```{r}
word_count_range1850s <- range(word_counts1850s)
word_count_range1950s <- range(word_counts1950s)
#1850s range 2396 17574
#1950s range 762 17321
```

```{r}
data <- data.frame(
  Year = c(1839, 1842, 1852, 1852, 1853, 1853, 1858, 1859, 1940, 1942,
           1945, 1948, 1950, 1950, 1950, 1951, 1952, 1952, 1954, 1954, 1955),
  ASL = c(32.958716, 24.089109, 9.452273, 28.935211, 28.892704, 20.823613, 26.641892, 
          31.801418, 12.713514, 12.625000, 21.357798, 14.830313, 9.020110, 19.818182, 
          11.798595, 24.065574, 12.590164, 9.076132, 10.485348, 22.316456, 14.430213),
  AWL = c(4.581013, 4.498561, 3.907167, 4.320058, 4.607382, 4.348931, 4.229521, 
          4.417706, 4.331188, 4.116343, 3.994943, 3.942544, 4.325866, 3.859438, 
          4.184095, 3.960326, 3.658392, 4.101703, 4.006108, 4.400313, 3.929711)
)

data_1800s <- subset(data, Year >= 1839 & Year <= 1859)
data_1900s <- subset(data, Year >= 1940 & Year <= 1955)

avg_1800s <- colMeans(data_1800s[, c("ASL", "AWL")])
avg_1900s <- colMeans(data_1900s[, c("ASL", "AWL")])
```


```{r}
data <- data.frame(
  Decade = c("1850s", "1950s"),
  Text_Count = c(8, 13),
    Total_Word_Count = c(69749, 69935)
,
  Mean_Word_Count = c(8718.625, 5379.615),
  Word_Range = c("2396 - 17574", "762 - 17321"),
 Mean_Average_Sentence_Length = c(25.449367, 15.009800),
  Mean_Average_Word_Length = c(4.363792, 4.062382)
  )

kable(data, caption = "Word Count Data for 1850s and 1950s", align = 'c', 
      col.names = c("Decade", "Texts", "Total Words", "Mean Words", "Word Range", "Average Sentence Length", "Average Word Length"))
```

# Methods

# Results

```{r}
# Calculate lexical density for each text
total_words <- ntoken(tokens_list)
lexical_words <- ntoken(tokens_lexical)
lexical_density <- lexical_words / total_words

lexical_density_data <- data.frame(
  Decade = c("1850s", "1950s"),
  Lexical_Density = c(0.481744825, 0.4976068)
)

kable(lexical_density_data, caption = "Average Lexical Density for 1850s and 1950s")

#0.481744825 is the average for 1850s
#0.4976068 is the average for 1950s
```

```{r}
lexdiv_ss_18 <- textstat_lexdiv(dfm_short_stories_18, measure = c("CTTR", "U", "S", "I", "Mass"))

lexdiv_ss_19 <- textstat_lexdiv(dfm_short_stories_19, measure = c("CTTR", "U", "S", "I", "Mass"))

### CTTR is the Corrected Type Token Ratio --> tells us lexical diversity corrected for text length

### U is the Uber Index --> this is another measure of lexical richness and diversity which is less sensitive to text length

### S is the Summers Index -> ????

### I is Yule's I --> quantifies the unevenness in the frequency distribution of words in a text; focuses on the distribution of unique words' frequencies


lex_div_res <- data.frame(
  Decade = c("1850s", "1950s"),
  CTTR = c( round(mean(lexdiv_ss_18$CTTR), 3),
                   round(mean(lexdiv_ss_19$CTTR), 3)
  ),
  Uber_Index = c(round(mean(lexdiv_ss_18$U), 3),
                round(mean(lexdiv_ss_19$U), 3)
  ),
  Summer_Index = c(round(mean(lexdiv_ss_18$S), 3),
                round(mean(lexdiv_ss_19$S), 3)
  ),
  Yule_index = c(round(mean(lexdiv_ss_18$I), 3),
                round(mean(lexdiv_ss_19$I), 3)
  )
)

kable(lex_div_res, caption = "Average Lexical for 1850s and 1950s", colnames = c("Decade", "Uber's Index", "Summer's Index", "Yule's Index"))

```

# Discussion

A future analysis could include more time periods and a larger variety of texts to improve the generalizability of our findings. Focusing on specific genres, such as romance or science-fiction, would allow for a better understanding of how different types of writing have evolved over time. This would help us understand whether certain genres have been more resistant to changes in style or if they have had the same characteristics over time. Additionally, analyzing how historical events or societal shifts influence language and writing styles could provide insight into how external factors influence the way people communicate.

# Acknowledgments {.appendix}

# Works Cited
1. Brown, D. W. (n.d.). textstat documentation. https://browndw.github.io/textstat_docs/
