---
title: "CBE 2"
author: "Amalia, Lauren, and Jenny"
date: last-modified
format:
  pdf:
    number-sections: true
    indent: true
    toc: true
    include-in-header:
      text: |
        \addtokomafont{disposition}{\rmfamily}
        \usepackage{fvextra}
        \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
execute:
  echo: false
---

\newpage

# Introduction

This project explores the differences in writing styles between the 1850s and the 1950s, with a focus on why older texts are often harder to understand. By examining a collection of short stories from both periods, we look at how factors like sentence length and lexical diversity may influence the readability of older texts. This question is interesting because understanding these shifts in writing style can reveal how language and communication have changed over time, and whether these shifts have made texts more accessible or harder to read.

# Data

```{r, include=FALSE}
library(quanteda)
library(quanteda.textstats)
library(tidyverse)
library(sentimentr)
library(knitr)
library(gt)
library(syuzhet)
library(tidytext)
library(dplyr)
source("../R/helper_functions.R")
source("../R/keyness_functions.R")

```

```{r}

txt_1945 <- paste(readLines("Text Files/the_other_side_of_the_hedge_1945.txt"), collapse = " ")

txt_1852 <- paste(readLines("Text Files/the_old_nurses_story_1852.txt"), collapse = " ")

txt_1952 <- paste(readLines("Text Files/the_old_man_at_the_bridge_1952.txt"), collapse = " ")

txt_1950 <- paste(readLines("Text Files/the_mindworm_1950.txt"), collapse = " ")

txt_1842 <- paste(readLines("Text Files/the_masque_of_the_red_death_1842.txt"), collapse = " ")

txt_1940 <- paste(readLines("Text Files/the_man_who_killed_the_world_1940.txt"), collapse = " ")

txt_1859 <- paste(readLines("Text Files/the_lifted_veil_1859.txt"), collapse = " ")

txt_1839 <- paste(readLines("Text Files/the_fall_of_the_house_of_usher_1839.txt"), collapse = " ")

txt_1852 <- paste(readLines("Text Files/the_district_doctor_1852.txt"), collapse = " ")

txt_1954 <- paste(readLines("Text Files/the_destructors_1954.txt"), collapse = " ")

txt_1942 <- paste(readLines("Text Files/the_catbird_seat_1942.txt"), collapse = " ")

txt_1858 <- paste(readLines("Text Files/psyche's_art_1858.txt"), collapse = " ")


txt_1954b <- paste(readLines("Text Files/master_zacharius_1954.txt"), collapse = " ")

txt_1955 <- paste(readLines("Text Files/good_country_people_1955.txt"), collapse = " ")


txt_1853 <- paste(readLines("Text Files/gods_in_exile_1853.txt"), collapse = " ")


txt_1950b <- paste(readLines("Text Files/first_confession_1950.txt"), collapse = " ")

txt_1950c <- paste(readLines("Text Files/coming_attraction_1950.txt"), collapse = " ")

txt_1853b <- paste(readLines("Text Files/bartelby_the_scrivener_1853.txt"), collapse = " ")

txt_1852b <- paste(readLines("Text Files/a_terribly_strange_bed_1852.txt"), collapse = " ")

txt_1952b <- paste(readLines("Text Files/a_sound_of_thunder_1952.txt"), collapse = " ")

txt_1948 <- paste(readLines("Text Files/a_perfect_day_for_bananafish_1948.txt"), collapse = " ")

txt_1951 <- paste(readLines("Text Files/A_&_P_1951.txt"), collapse = " ")

```

```{r,include=FALSE}
years_q <- c("1839", "1842", "1852", "1852b", "1853", "1853b", "1858", "1859", "1940", "1942", "1945", "1948", "1950", "1950b", "1950c", "1951", "1952", "1952b", "1954", "1954b", "1955")

years <- c(txt_1839, txt_1842, txt_1852, txt_1852b, txt_1853, txt_1853b, txt_1858, txt_1859, txt_1940, txt_1942, txt_1945, txt_1948, txt_1950, txt_1950b, txt_1950c, txt_1951, txt_1952, txt_1952b, txt_1954, txt_1954b, txt_1955)

years_18_q <- c("1839", "1842", "1852", "1852b", "1853", "1853b", "1858", "1859")
years_19_q <- c("1940", "1942", "1945", "1948", "1950", "1950b", "1950c", "1951", "1952", "1952b", "1954", "1954b", "1955")

years_18 <- c(txt_1839, txt_1842, txt_1852, txt_1852b, txt_1853, txt_1853b, txt_1858, txt_1859)
years_19 <- c(txt_1940, txt_1942, txt_1945, txt_1948, txt_1950, txt_1950b, txt_1950c, txt_1951, txt_1952, txt_1952b, txt_1954, txt_1954b, txt_1955)


short_stories_corpus <- data.frame(doc_id = years_q, text = years) %>%
   mutate(text = preprocess_text(text)) %>%
   corpus()

short_stories_18_corpus <- data.frame(doc_id = years_18_q, text = years_18) %>%
   mutate(text = preprocess_text(text)) %>%
   corpus()

short_stories_19_corpus <- data.frame(doc_id = years_19_q, text = years_19) %>%
   mutate(text = preprocess_text(text)) %>%
   corpus()

dfm_short_stories <- dfm_remove(dfm(tokens(short_stories_corpus)), stopwords("en"))

dfm_short_stories_18 <- dfm_remove(dfm(tokens(short_stories_18_corpus)), stopwords("en"))

dfm_short_stories_19 <- dfm_remove(dfm(tokens(short_stories_19_corpus)), stopwords("en"))
```

```{r}
tokens_list <- tokens(short_stories_corpus)
tokens_lexical <- tokens_select(tokens_list, pattern = stopwords("en"), selection = "remove")
```

```{r}
calculate_asl <- function(text) {
   unlist(str_split(text, pattern = "\\.\\s*")) %>%
    sapply(function(x) str_count(x, "\\w+")) %>%
    mean()
}


calculate_awl <- function(tokens) {
  tokens %>%
    tokens(what = "word") %>%
    as.list() %>%
    unlist() %>%
    nchar() %>%
    mean()
}

results <- tibble(
  Year = years_q,
  ASL = sapply(years, calculate_asl),
  AWL = sapply(short_stories_corpus, calculate_awl)
)

results %>%
  arrange(Year) %>%
  print()

```

```{r}
all_text <- paste(txt_1839, txt_1842, txt_1852, txt_1852b, txt_1853, txt_1853b, txt_1858, txt_1859)
words <- unlist(strsplit(all_text, "\\s+"))
total_word_count <- length(words)
print(total_word_count)
all_text1950s <- paste(txt_1940, txt_1942, txt_1945, txt_1948, txt_1950, txt_1950b, txt_1950c, txt_1951, txt_1952, txt_1952b, txt_1954, txt_1954b, txt_1955)
words1950s <- unlist(strsplit(all_text1950s, "\\s+"))
total_word_count1950s <- length(words1950s)
```
```{r}
word_count_1839 <- length(unlist(strsplit(txt_1839, "\\s+")))
word_count_1842 <- length(unlist(strsplit(txt_1842, "\\s+")))
word_count_1852 <- length(unlist(strsplit(txt_1852, "\\s+")))
word_count_1852b <- length(unlist(strsplit(txt_1852b, "\\s+")))
word_count_1853 <- length(unlist(strsplit(txt_1853, "\\s+")))
word_count_1853b <- length(unlist(strsplit(txt_1853b, "\\s+")))
word_count_1858 <- length(unlist(strsplit(txt_1858, "\\s+")))
word_count_1859 <- length(unlist(strsplit(txt_1859, "\\s+")))
word_counts1850s <- c(word_count_1839, word_count_1842, word_count_1852, word_count_1852b, word_count_1853, word_count_1853b, word_count_1858, word_count_1859)
average_word_count1850s <- mean(word_counts1850s)

#average word count 1850s: 8718.625
#average word count 1950s: 5379.615
```

```{r}
word_count_1940 <- length(unlist(strsplit(txt_1940, "\\s+")))
word_count_1942 <- length(unlist(strsplit(txt_1942, "\\s+")))
word_count_1945 <- length(unlist(strsplit(txt_1945, "\\s+")))
word_count_1948 <- length(unlist(strsplit(txt_1948, "\\s+")))
word_count_1950 <- length(unlist(strsplit(txt_1950, "\\s+")))
word_count_1950b <- length(unlist(strsplit(txt_1950b, "\\s+")))
word_count_1950c <- length(unlist(strsplit(txt_1950c, "\\s+")))
word_count_1951 <- length(unlist(strsplit(txt_1951, "\\s+")))
word_count_1952 <- length(unlist(strsplit(txt_1952, "\\s+")))
word_count_1952b <- length(unlist(strsplit(txt_1952b, "\\s+")))
word_count_1954 <- length(unlist(strsplit(txt_1954, "\\s+")))
word_count_1954b <- length(unlist(strsplit(txt_1954b, "\\s+")))
word_count_1955 <- length(unlist(strsplit(txt_1955, "\\s+")))

word_counts1950s <- c(word_count_1940, word_count_1942, word_count_1945, word_count_1948, word_count_1950, word_count_1950b, word_count_1950c, 
                      word_count_1951, word_count_1952, word_count_1952b, 
                      word_count_1954, word_count_1954b, word_count_1955)
average_word_count1850s <- mean(word_counts1950s)
```

```{r}
word_count_range1850s <- range(word_counts1850s)
word_count_range1950s <- range(word_counts1950s)
#1850s range 2396 17574
#1950s range 762 17321
```

```{r}
data <- data.frame(
  Year = c(1839, 1842, 1852, 1852, 1853, 1853, 1858, 1859, 1940, 1942,
           1945, 1948, 1950, 1950, 1950, 1951, 1952, 1952, 1954, 1954, 1955),
  ASL = c(32.958716, 24.089109, 9.452273, 28.935211, 28.892704, 20.823613, 26.641892, 
          31.801418, 12.713514, 12.625000, 21.357798, 14.830313, 9.020110, 19.818182, 
          11.798595, 24.065574, 12.590164, 9.076132, 10.485348, 22.316456, 14.430213),
  AWL = c(4.581013, 4.498561, 3.907167, 4.320058, 4.607382, 4.348931, 4.229521, 
          4.417706, 4.331188, 4.116343, 3.994943, 3.942544, 4.325866, 3.859438, 
          4.184095, 3.960326, 3.658392, 4.101703, 4.006108, 4.400313, 3.929711)
)

data_1800s <- subset(data, Year >= 1839 & Year <= 1859)
data_1900s <- subset(data, Year >= 1940 & Year <= 1955)

avg_1800s <- colMeans(data_1800s[, c("ASL", "AWL")])
avg_1900s <- colMeans(data_1900s[, c("ASL", "AWL")])
```


```{r}
data <- data.frame(
  Decade = c("1850s", "1950s"),
  Text_Count = c(8, 13),
    Total_Word_Count = c(69749, 69935)
,
  Mean_Word_Count = c(8718.625, 5379.615),
  Word_Range = c("2396 - 17574", "762 - 17321"),
 Mean_Average_Sentence_Length = c(25.449367, 15.009800),
  Mean_Average_Word_Length = c(4.363792, 4.062382)
  )

kable(data, caption = "Word Count Data for 1850s and 1950s", align = 'c', 
      col.names = c("Decade", "Texts", "Total Words", "Mean Words", "Word Range", "Average Sentence Length", "Average Word Length"))
```

# Methods

For our initial analysis, we wanted to determine the complexity of the texts from the different time periods. In order to do so, we analyzed two different common measurements for lexical complexity - lexical density and lexical diversity (CITATION HERE FOR ARTICLE). 

First we examined the lexical density of texts from the 1850s and 1950s. Lexical density is a measure of the proportion of content words to function words in a text. Content words, which include nouns, verbs, adjectives, etc, are words that carry meaning in a sentence. On the other hand, function words, such as articles, conjunctions, prepositions, etc, are used for grammatical purposes. Lexical density is higher when the proportion of content words is greater than that of function words. 

We then examined lexical diversity using two different measurements - Carroll's Corrected Type Token Ratio (CTTR), Uber's Index, Summer's Index, and Yule's K. We used three different measurements to provide confirmation on the diversity analysis. These three measurements were chosen because they all account and correct for text length sensitivities. Yule's K was chosen because it is sensitive to frequency distributions. A low score for all measurements typically indicates lower diversity when compared to a higher score. Yule's K provides additional information concerning the distribution of word frequencies, with a lower value typically indicating an even distribution.

# Results

```{r}
# Calculate lexical density for each text
total_words <- ntoken(tokens_list)
lexical_words <- ntoken(tokens_lexical)
lexical_density <- lexical_words / total_words

lexical_density_data <- data.frame(
  Decade = c("1850s", "1950s"),
  Lexical_Density = c(0.481744825, 0.4976068)
)

colnames(lexical_density_data) <- c("Decade", "Lexical Diversity")

kable(lexical_density_data, caption = "Average Lexical Density for 1850s and 1950s")

#0.481744825 is the average for 1850s
#0.4976068 is the average for 1950s
```

The average lexical density for the texts from the 1850s is 0.4817, while for the 1950s, it is 0.4976. This indicates a small increase in lexical density over time, suggesting a slight rise in the proportion of content words in the 1950s texts. However, given the minimal difference between the two decades, this change is unlikely to have a significant impact on the readability or complexity of the texts. A larger dataset would be needed to draw more definitive conclusions, since our sample size is relatively small.

```{r}
lexdiv_ss_18 <- textstat_lexdiv(dfm_short_stories_18, measure = c("CTTR", "U", "S", "I", "Mass"))

lexdiv_ss_19 <- textstat_lexdiv(dfm_short_stories_19, measure = c("CTTR", "U", "S", "I", "Mass"))

### CTTR is the Corrected Type Token Ratio --> tells us lexical diversity corrected for text length

### U is the Uber Index --> this is another measure of lexical richness and diversity which is less sensitive to text length

### S is the Summers Index -> measures the lexical richness of vocabulary in a text

### I is Yule's I --> quantifies the unevenness in the frequency distribution of words in a text; focuses on the distribution of unique words' frequencies


lex_div_res <- data.frame(
  Decade = c("1850s", "1950s"),
  CTTR = c( round(mean(lexdiv_ss_18$CTTR), 3),
                   round(mean(lexdiv_ss_19$CTTR), 3)
  ),
  Uber_Index = c(round(mean(lexdiv_ss_18$U), 3),
                round(mean(lexdiv_ss_19$U), 3)
  ),
  Summer_Index = c(round(mean(lexdiv_ss_18$S), 3),
                round(mean(lexdiv_ss_19$S), 3)
  ),
  Yule_index = c(round(mean(lexdiv_ss_18$I), 3),
                round(mean(lexdiv_ss_19$I), 3)
  )
)

colnames(lex_div_res) <- c("Decade", "CTTR", "Uber's Index","Summer's Index", "Yule's Index")


kable(lex_div_res, caption = "Average Lexical Diversity for 1850s and 1950s")


```

For the CTTR, Uber's Index, and Summer's Index, the average lexical diversity of the texts from the 1850s is higher than the average lexical diversity from the 1950s. This indicates a decrease in lexical diversity between the two time periods. Conversely, the Yule's K value for the 1850s is 193.724 where the Yule's K value for the 1950s is 98.966. This suggests that the texts from the 1950s have a more uniform distribution of word frequencies and thus less lexical diversity.

# Discussion

A future analysis could include more time periods and a larger variety of texts to improve the generalizability of our findings. Focusing on specific genres, such as romance or science-fiction, would allow for a better understanding of how different types of writing have evolved over time. This would help us understand whether certain genres have been more resistant to changes in style or if they have had the same characteristics over time. Additionally, analyzing how historical events or societal shifts influence language and writing styles could provide insight into how external factors influence the way people communicate.

# Acknowledgments {.appendix}

# Works Cited
1. Brown, D. W. (n.d.). textstat documentation. https://browndw.github.io/textstat_docs/
